Precise spiking motifs in neurobiological and neuromorphic data

### 2022-11-08 manuscript.tex
Update on Overleaf.

More recently, novel multi-unit recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}. Authors review these findings, consider the mechanisms by which travelling waves are generated, and evaluate their possible roles in cortical function. In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{davis_spontaneous_2021}. Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally-generated noise. However, fluctuations of spontaneous activity, often organized as traveling waves, shape stimulus-evoked responses and perceptual sensitivity. The mechanisms underlying these waves are unknown. Further, it is unclear whether waves are consistent with the low rate and weakly correlated ``asynchronous-irregular'' dynamics observed in cortical recordings. Here, the authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales. They find that spontaneous traveling waves are a general property of these networks. The traveling waves that occur in the model are sparse, with only a small fraction of neurons participating in any individual wave. Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states. Further, by modulating local network state, they can shape responses to incoming inputs as observed \textit{in vivo}. Such waves also occur in motor areas and~\citet{linden_movement_2021} have presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.

>>>>

More recently, novel multi-unit recording techniques have enabled the identification of travelling waves of neural activity in different areas of the cortex~\citep{muller_cortical_2018}. Authors review these findings by considering the mechanisms by which travelling waves are generated, and evaluate their possible roles in cortical function. In particular, spontaneous traveling waves naturally emerge from horizontal fiber time delays and travel through locally asynchronous-irregular states~\citep{davis_spontaneous_2021}. Studies of sensory-evoked neuronal responses often focus on mean spike rates, with fluctuations treated as internally-generated noise. However, fluctuations of spontaneous activity, often organized as traveling waves, shape stimulus-evoked responses and perceptual sensitivity. The mechanisms underlying these waves are unknown. Further, it is unclear whether waves are consistent with the low rate and weakly correlated ``asynchronous-irregular'' dynamics observed in cortical recordings. In that paper, the authors describe a large-scale computational model with topographically-organized connectivity and conduction delays relevant to biological scales. They find that spontaneous traveling waves are a general property of these networks. The traveling waves that occur in the model are sparse, with only a small fraction of neurons participating in any individual wave. Consequently, they do not induce measurable spike correlations and remain consistent with locally asynchronous irregular states. Further, by modulating the state of the local network, they can shape responses to incoming inputs as observed \textit{in vivo}. Such waves also occur in motor areas and~\citet{linden_movement_2021} have presented ensemble-recordings of neurons in the lumbar spinal cord that indicate that, rather than alternating, the population is performing a low-dimensional ``rotation'' in neural space, in which the neural activity is cycling through all phases continuously during the rhythmic behavior.
